### 2022-12-14 lmcs/REBUTTAL-LETTER.md

- P6, grammar, bottom of page: Why is K (defined earlier in the term syntax) a value? Shouldn’t K be mentioned in rhs of E? Also, I do not understand the need for the thread evaluation context, nor do I understand how it fits with the remaining notions, E and G. F doesn’t show up there. Could it be that F should appear in the rhs of G? Maybe instead of the hole?~~

  The authors: Functions are values, K is the sort for builtin functions, so K is a value. The thread evaluation context is separately useful. The contexts F and G have holes of different types. They are used in different places.

- SC-ResComm: Why is the side condition necessary? Doesn’t name restriction generate fresh names?

  The authors: Yes, operationally, the `new` function generates name restrictions with fresh names. However, this is not encoded as a property of the _syntax_ of terms. Therefore, to ensure that the equivalence works for any configuration, and not just configurations which are reachable by the operational semantics, it needs that side-condition.

@@

- P8, ”We no longer require that every child thread finishes by returning a terminated channel”: Doesn’t that have implications for linearity?

  The authors: It doesn't. In previous settings, each child thread returned its terminated channel to be closed by the "runtime system". In our setting, it is enforced by the type system that each child thread must close its terminated channel.

- Fig2: rule T-Absurd would admit weakening, so is wrong.

  The authors: The rule T-Absurd is exfalso, i.e., from a proof of false, or a term of the empty type, anything follows, including weakening.

  We believe that T-Absurd can only be used to derive weakening in the presence of a proof of false. If the reviewer believes that T-Absurd admits general weakening, we would welcome them to elaborate.

@@

- T-Pair: why should p < minpr(∆)?

  The authors: We need this condition because of the operational semantics evaluating products, which evaluate the fields of a pair in order.

@@

  The authors: Lemma 3.3 is subject reduction for a simply-typed lambda calculus with constants, which is standard. We clarified this in the paper.

- Also, I am not familiar with type schemas, some explanation would be helpful.

  The authors: we have now fixed use of A,B,C in type schemas and added further explanation.

  The authors: This is the composition of the structural congruence and the reduction relation. We split them up and elaborated in the text:

  > In GV, this is not the case, and subject reduction relies on proving that if the result of rewriting via ≡ followed by reducing via −→C is an ill-typed configuration, we can rewrite it to a well-typed configuration via ≡.

@@

We have fixed all the minor comments below.

- We have fixed all occurances of bare `new`.

- P2, ”communicate via exactly one series of channels”: What is meant by series of channels?

- P28, ”Let this process be let this be Pi...”: grammar

- P36: It would be helpful to already alert the reader to the figure on the next page when discussing the setup here.

@@

- P3, ”is typeable and guaranteed deadlock-free”: Guaranteed to be

- P5: ”to denote the minimum and maximum”: ..., respectively. Also, that listing seems to be missing max!

  The authors: we defined max here as we need it and it is used later.

- P5: The programs in Example 2.1 are not well-formed wrt abstract syntax of terms defined on page 5. For example, the term ”recv x” cannot be parsed.

  The authors: The term `recv x` is well-formed. Specifically, it is an application (of the form `M N`) where:
    * `M` is a constant (of the form `K`), specifically, it is `recv`; and
    * `N` is a variable (of the form `x`).

### 2022-12-14 lmcs/priority-gv.tex
No more TODO's in REBUTTAL_LETTER

While our configuration reduction is based on the standard evaluation for GV, the increased expressiveness of PGV allows us to simplify the relation on two counts.
\begin{enumerate}[label=(\alph*)]
\item
\emph{We decompose the $\tm{\fork}$ construct}.
In GV, $\tm{\fork}$ creates a new channel, spawns a child thread, and, when the child thread finishes, it closes the channel to its parent. In PGV, these are three separate operations: $\tm{\new}$, $\tm{\spawn}$, and $\tm{\close}$. We no longer require that every child thread finishes by returning a terminated channel. Consequently, we also simplify the evaluation of the $\tm{\link}$ construct.

Intuitively, evaluating $\tm{\link}$ causes a substitution: if we have a channel bound as $\tm{\res{x}{y}{}}$, then $\tm{\link\;\pair{w}{x}}$ replaces all occurrences of $\tm{y}$ by $\tm{w}$. However, in GV, $\tm{\link}$ is required to return a terminated channel, which means that the semantics for $\tm{link}$ must {create} a fresh channel of type $\ty{\tyends}/\ty{\tyendr}$. The endpoint of type $\ty{\tyends}$ is returned by the $\tm{link}$ construct, and a $\tm{\wait}$ on the other endpoint guards the {actual} substitution. In PGV, evaluating $\tm{\link}$ simply causes a substitution.
\item
\emph{Our structural congruence is type preserving}. Consequently, we can embed it directly into the reduction relation. In GV, this is not the case, and subject reduction relies on proving that if $\equiv\cred$ ends up in an ill-typed configuration, we can rewrite it to a well-typed configuration using $\equiv$.
\end{enumerate}

@@

\begin{lemma}[Progress, $\tred$]%
  \label{lem:pgv-open-progress-terms}
  If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ and $\ty{\Gamma}$ contains only session types, then:
  \begin{enumerate*}[label= (\alph*) ]
  \item $\tm{M}$ is a value;
  \item $\tm{M}\tred\tm{N}$ for some $\tm{N}$; or
  \item $\tm{M}$ is ready.
  \end{enumerate*}
\end{lemma}

>>>>

While our configuration reduction is based on the standard evaluation for GV, the increased expressiveness of PGV allows us to simplify the relation on two counts.
\begin{enumerate}[label=(\roman*)]
\item
\emph{We decompose the $\tm{\fork}$ construct}.
In GV, $\tm{\fork}$ creates a new channel, spawns a child thread, and, when the child thread finishes, it closes the channel to its parent. In PGV, these are three separate operations: $\tm{\new}$, $\tm{\spawn}$, and $\tm{\close}$. We no longer require that every child thread finishes by returning a terminated channel. Consequently, we also simplify the evaluation of the $\tm{\link}$ construct.

Intuitively, evaluating $\tm{\link}$ causes a substitution: if we have a channel bound as $\tm{\res{x}{y}{}}$, then $\tm{\link\;\pair{w}{x}}$ replaces all occurrences of $\tm{y}$ by $\tm{w}$. However, in GV, $\tm{\link}$ is required to return a terminated channel, which means that the semantics for $\tm{link}$ must {create} a fresh channel of type $\ty{\tyends}/\ty{\tyendr}$. The endpoint of type $\ty{\tyends}$ is returned by the $\tm{link}$ construct, and a $\tm{\wait}$ on the other endpoint guards the {actual} substitution. In PGV, evaluating $\tm{\link}$ simply causes a substitution.
\item
\emph{Our structural congruence is type preserving}. Consequently, we can embed it directly into the reduction relation. In GV, this is not the case, and subject reduction relies on proving that if the result of rewriting via $\equiv$ followed by reducing via $\cred$ is an ill-typed configuration, we can rewrite it to a well-typed configuration via $\equiv$.
\end{enumerate}

@@

\begin{lemma}[Progress, $\tred$]%
  \label{lem:pgv-open-progress-terms}
  If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ and $\ty{\Gamma}$ contains only session types, then:
  \begin{enumerate*}[label= (\roman*) ]
  \item $\tm{M}$ is a value;
  \item $\tm{M}\tred\tm{N}$ for some $\tm{N}$; or
  \item $\tm{M}$ is ready.
  \end{enumerate*}
\end{lemma}
